### 2013-07-07 paper.tex
Updated and reworked parts according to the prvious comments.

\section{Introduction}
Tree volume measurement is one of the many parameters in documenting and 
profiling trees and is at the center of forest inventory. The use of this 
information is important to companies who need to make smart, profitable 
decisions but also need to be certain that the choices they make now are 
maintable in the future. However this information also benefits climate 
modellers, environmental groups and governments who want to be able to sustain 
ecosystems and plan restorations. To be able to accomplish this in forestry, 
it is a necessity to be able to collect large amounts of data easily and on a 
regular basis. Most methods used today to measure individual trees are often 
done manually with a variety of tools. The use of these analogue approaches 
is expensive, time consuming and in most cases the data collected doesn't meet 
today's standards \cite{digital imaged based tree measurement for forest inventory}. 
SilviaTerra, however, is a company willing to change that and have already 
developed several tools for scanning, analyzing and mapping areas of forestry 
to give accurate results. This is achieved through computer exactitude combined 
with statistical analysis. Recently a new project was started, with this 
research paper developing the mathematical back end, to allow users to easily 
get a correct volume measurement by only using their smart phone.

The measurements required for forest inventory can be achieved through a number 
of methods. One of them is using a so called biltmore stick. A biltmore stick 
is similar to a normal ruler but has been specially modified for measuring trees. 
When standing at a certain distance the user can discern the height of the tree 
by holding the biltmore stick up to their eye and reading the corresponding number. 
The biltmore stick can also easily be used to measure the DBH of the tree. 
Other ways of measuring tree volume include using a combination of lasers and 
inclinometers or by having a person manually climb up the tree and drop a tape 
measure down to measure the height. Despite these relatively simpe ways of 
measuring they often leave room for error. That’s why a more common way of 
getting a precise reading is to cut down trees in the area and then physically 
measure them. After the relevant data is collected one applies a stem-profile 
taper equation that is specific to the species of tree measured. These taper 
functions are based on numerous measurements and have been refined over the years. 
The main problem with taper functions is that to get an accurate result they 
need to be callibrated to the hundreds of local conditions. This again is cost 
prohibitive and only so much of a forest can be accounted for in the taper function. 
Even when one manages to create a new taper function, it is still far from 
perfect as all trees, even from the same species, differ in shape.

This project aims to solve the described problem by using a combination of 
photos taken of the tree, various sensor data from the phone and mathematical 
computation. This is supposed to then give a accurate result, cheaply and 
non-destructively. All within the reaches of a mobile phone.

@@

After that they take two pictures. The interface is quite simple with a 
cross-hair to show where to point the camera. The first picture is of the 
actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, 
figure \ref{bole}. Between these two pictures the camera will have to be tilted 
vertically to capture the bole. The angle between these two camera positions is 
recorded through the internal gyroscope. After this stage the user's role in 
the application is done and the pictures is sent to one of SilviaTerra's web 
servers. Here horizontal lines that run parallel with the bottom of the image 
are drawn across the entire picture. These are drawn at different pixel heights 
with equal pixel distance in between. The reason this is done is because 
computer vision is still an experimental technology and is not fully capable 
yet of being to detect the edges of trees with a background of forestry. 
Seeing as this is a commercial product, it would be quite disadvantageous if 
errors were commonplace. So instead the images with the newly drawn lines are 
then forwarded to Amazon's Mechanical Turk service. Just like the name suggests 
Amazon's Mechanical Turk is meant to be analogy to the 18th century concept of 
a automated chess board. The service is in its simplest form a crowd sourcing 
internet marketplace where individuals or businesses can request tasks to be 
completed. Usually the tasks are ones that are difficult for computers but 
easy for humans. SilviaTerra creates tasks where people manually click where 
the previously drawn lines, cross the edges of the tree. Ones of the points of 
intersection are identified, their corresponding pixel coordinates are the sent 
back to the web server. The pixel coordinates along with information from the 
phone is piped into a program where the volume is then calculated. It is this 
program that is the main topic of this research paper.

@@

\subsection{Distance to tree}
The DBH is used as a reference to find all the other measurements. The first
unknown that must be discovered is the distance to the tree. This distance is 
parallel with the ground. The difference between the x-coordinates of the left 
and right pixel that correspond to the DBH is calculated, figure 
\ref{horizontal_triangle}, $X_1$ and $X_2$. Due to the fact that the width of
the DBH is now known in both pixels and in a physical measurement, the ratio
between the two can be determined by dividing the physical distance between 
$X_1$ and $X_2$ with the number of pixels betwee them.
\begin{figure}[htp]
\centering
{
	\includegraphics[scale=0.3]{horizontal_triangle.pdf}
	\caption{Birds eye view of horizontal plane}
	\label{horizontal_triangle}
}
\end{figure}
 As seen in figure \ref{horizontal_triangle} an isosceles triangle can be 
 constructed with the horizontal angle of view as the vertex angle. The triangles 
 two congruent legs extend out towards the base of the triangle, $W_1$, which 
 is inline with the tree. To get the length of the base, $W_1$, the previously 
 calculated scale between pixels and the real world is multiplied with the 
 number of pixels in the horizontal resolution of the image. The distance to 
 the tree can easily calculated by drawing an extra line that bisects the vertex
 angle, contriving the angle $\theta$ which is exactly half of the angle of view 
 and the actual distance to the tree $d$. This line also divides $W_1$ into $W_2$ 
 and futhermore producing a right angle triangle that can be solved. Finally to 
 actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.

@@

The next stage is retrieving the real world height to each set pixel coordinates 
as well as the width between them, which will be the diameter of the tree. The 
first step taken is to get the height. The vertical image plate scale which is 
the ratio between degrees and pixels in the vertical axis is calculated by 
dividing the vertical resolution of the image with the vertical angle of view. 
This can be done because it is assumed that the silhouette of the tree is on a
plane in the real world which corresponds to the image plane. With this method 
it is possible to find out the angle to each pixel coordinate, $P$ in figure 
\ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is 
denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However 
if the pixel coordinate was identified in the second image, when the phone was 
tilted upwards, then angle between the distance to the tree and the lower boundary
of the vertical angle of view, $B$, will need to be added, which is represented
by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also 
calculated so that the width can later be determined.

@@

\subsection{Volume}
Now the only thing left to produce the final result is to use these newly 
calculated widths and heights to obtain the total volume. The current implemented 
method of doing this is by integrating all the points using the trapezoidal rule.
All the widths are divided by 2 to get the radius. Then they are squared and
multiplied with $\pi$ to the obtain the basal area at each height. The basal 
area is plotted along with their corresponding heights. The trapezoidal rule 
then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$.
The area between all the points is the summed together.

@@

\subsection{Simulation}
To test this method a \emph{virtual tree} was constructed. This tree was made 
out of blue tape and tapered off so that the program could be thoroughly tested. 
Along the tape points were marked out and the height to them plus the width
between corresponding points were measured. Just like in the real application, 
two pictures were taken. One of the DBH and one of the top of the tree. Each 
set of points was then identified on the pictures and measured in pixels. The 
parameters were then passed into the program and the results were cross-referenced
with the real life measurements.

@@

\begin{table}[h!]
	\begin{center}
    	\begin{tabular}{| l c c c c c c r |}
    	\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		412222   & 386642    & 396386    & 418370    & 412084    & 433294    & 445401    & 445624    \\
		\hline
		\end{tabular}
		\caption{Total volume over DBH}
		\label{volumes}
    \end{center}
\end{table}

\section{Discussion}
\subsection{Error tolerance}
As seen in table \ref{heights} either when the vertical tilt angle is very wide, 
standing close to the tree, or very narrow, standing far away, leads to error. I
t's not clear what is causing this but there a two possibl reasons. The first one is
that because of the short focal length of the camera and overall poor quality when 
compared to a real camera, causes various forms of distrotion, especially around the 
edges of the image. When the person is standing close to the tree so the vertical tilt 
angle is over 45\degree, the points tend to lie exactly there, around the edges of 
the image. The distortion then affects these points causing the pixel measurements 
to be incorrect. The other problem occurs in the direct opposite situation, when 
the person is standing far away and the angle is very narrow. Naturally what happens 
is that the resolution of the camera just isn't high enough and the pixel measurements 
just aren't accurate anymore. However most of this distortion seems to only affect the 
vertical measurements because when looking at table \ref{widths} it is clearly seen that 
the readings do not stray much at all from their actual values. Despite this the deviation 
in the height readings still alters the volume calculations.
\subsection{Limitations}
\subsection{Future development}
\subsection{Future development}

>>>>

\section{Introduction}
Tree volume measurement is one of the many parameters in documenting and profiling trees and is at the center of forest inventory.Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration. The need for information about Earth's natural resources, including tree volume, has been increasing exponentially in recent times. To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis.

Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}
Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches do not come without their downfalls.

The measurements required for forest inventory can be achieved through a number of methods. One of them is using a so called biltmore stick. A biltmore stick is similar to a normal ruler but has been specially modified for measuring trees. When standing at a certain distance the user can discern the height of the tree by holding the biltmore stick up to their eye and reading the corresponding number. The biltmore stick can also easily be used to measure the \emph{DBH} (Diameter at Breast Height) of the tree. Of course a biltmore stick tries to be as generic as possible so that it can be used on a variety of trees and therefore has no guarantee of providing an accurate result. Other ways of measuring tree volume include using a combination of lasers and inclinometers or by having a person manually climb up the tree and drop a tape measure down to measure the height. Despite these relatively simple ways of measuring they often leave room for error. That’s why a more common way of getting a precise reading is to cut down trees in the area and then physically measuring them. After the relevant data is collected one applies a stem-profile taper equation that is specific to the species of tree measured. These taper functions are based on numerous measurements and have been refined over the years. The main problem with taper functions is that to get an accurate result they need to be callibrated to the hundreds of local conditions and are usually custom made for a certain species in a certain area. This again is cost prohibitive and only so much of a forest can be accounted for in the taper function. Even when one manages to create a new taper function, it is still far from perfect as all trees, even from the same species, differ in shape.

SilviaTerra, a high-tech startup company, however is willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results. This is achieved through computer exactitude combined with statistical analysis. Recently a new project was started, with this research project developing the mathematical back end, to allow users to easily get a correct volume measurement by only using their smart phone.This project aims to solve the problem of obtaining an accurate volume reading of tree, cheaply and non-destructively. This will be done by using a combination of photos taken of the tree, various sensor data from the phone and mathematical computation.

@@

After that they take two pictures. The interface is quite simple with a cross-hair to show where to point the camera. The first picture is of the actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, figure \ref{bole}. Between these two pictures the camera will have to be tilted vertically to capture the bole. The angle between these two camera positions is recorded through the internal gyroscope and accelerometer. After this stage the user's role in the application is done. Now the challenge is to extract useful information from these images. The pictures are sent to one of SilviaTerra's web servers. Here horizontal lines that run parallel with the bottom of the image are drawn across the entire picture. These are drawn at different pixel heights with equal pixel distance in between. The reason this is done is because computer vision is still an experimental technology and is computationally difficult as well as often being error-prone. SilviaTerra has instead opted a different route where the task of edge detection is performed manually by workers at Amazon's Mechanical Turk. So the images with the newly drawn lines are then forwarded to Mechanical Turk. The service is in its simplest form a crowd sourcing internet marketplace where individuals or businesses can request tasks to be completed. Usually the tasks are ones that are difficult for computers but easy for humans. SilviaTerra creates tasks where people manually click where the previously drawn lines, cross the edges of the tree. Once the points of intersection are identified, their corresponding pixel coordinates are then paired with the phone sensor data and undergo photogrammetric analysis.

@@

\subsection{Distance to tree}
The DBH is used as a reference to find all the other measurements. The first unknown that must be discovered is the distance to the tree. This distance is parallel with the ground. The difference between the x-coordinates of the left and right pixel that correspond to the DBH is calculated, figure \ref{horizontal_triangle}, $X_1$ and $X_2$. Due to the fact that the width of the DBH is now known in both pixels and in a physical measurement, the ratio between the two can be determined by dividing the physical distance between $X_1$ and $X_2$ with the number of pixels between them. For example if the width of the tree is 100 pixels and its real life width is 50 centimeters then ratio is 0.5 cm per pixel.
\begin{figure}[htp]
\centering
{
	\includegraphics[scale=0.3]{horizontal_triangle.pdf}
	\caption{Birds eye view of horizontal plane}
	\label{horizontal_triangle}
}
\end{figure}
As seen in figure \ref{horizontal_triangle} an isosceles triangle can be constructed with the horizontal angle of view as the vertex angle. The triangles two congruent legs extend out towards the base of the triangle, $W_1$, which is inline with the tree. To get the length of the base, $W_1$, the previously calculated scale between pixels and the real world is multiplied with the number of pixels in the horizontal resolution of the image. The distance to the tree can easily calculated by drawing an extra line that bisects the vertex angle, resulting in an angle, $\theta$, which is exactly half of the angle of view and the actual distance to the tree $d$. This line also divides $W_1$ into $W_2$ and futhermore produces a right angle triangle that can be solved. Finally to actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.

@@

The next stage is retrieving the real world height to each set pixel coordinates as well as the width between them, which will be the diameter of the tree. The first step taken is to get the height. The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view. This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor. With this method it is possible to find out the angle to each pixel coordinate, example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also calculated so that the width can later be determined.

@@

\subsection{Volume}
Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume. The current implemented method of doing this is by integrating all the points using the trapezoidal rule. All the widths are divided by 2 to get the radius. Then they are squared and multiplied with $\pi$ to the obtain the basal area at each height. The basal area is plotted along with their corresponding heights. The trapezoidal rule then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$. The area between all the points is then summed together.

@@

\subsection{Simulation}
To test this method a \emph{virtual tree}, that represented a 4m tall tree,was constructed. This tree was made out of blue tape and tapered off drastically so that the program could be thoroughly tested. Along the tape, points were marked out and the height to them plus the width between corresponding points was measured. Just like in the real application, two pictures were taken. One of the DBH and one of the top of the tree. Each set of points was then identified on the pictures and measured in pixels. The parameters were then passed into the program and the results were cross-referenced with the real life measurements.

@@

\begin{table}[h!]
	\begin{center}
    	\begin{tabular}{| l c c c c c c r |}
    	\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		452095   & 418053    & 428738    & 455700    & 451214    & 481471    & 501534    & 492573    \\
		\hline
		\end{tabular}
		\caption{Total volume over DBH}
		\label{volumes}
    \end{center}
\end{table}

\section{Discussion}
\subsection{Error tolerance}
As seen in table \ref{heights} either when the vertical tilt angle is very wide, standing close to the tree, or very narrow, standing far away, leads to error. It's not clear what is causing this but there a two possibl reasons. The first one is that because of the short focal length of the camera and overall poor quality when compared to a real camera, causes various forms of distrotion, especially around the edges of the image. When the person is standing close to the tree so the vertical tilt angle is over 45\degree, the points tend to lie exactly there, around the edges of the image. The distortion then affects these points causing the pixel measurements to be incorrect. The other problem occurs in the direct opposite situation, when the person is standing far away and the angle is very narrow. Naturally what happens is that the resolution of the camera just isn't high enough and the pixel measurements just aren't accurate anymore. However most of this distortion seems to only affect the vertical measurements because when looking at table \ref{widths} it is clearly seen that the readings do not stray much at all from their actual values. Despite this the deviation in the height readings still alters the volume calculations. Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which is basically means they should stand somewhere between one tree length to two tree length's from the tree.
\subsection{Limitations}
As pointed out in the previous section the main hindrance attatched to this application is the actual camera. 
\subsection{Future development}
The main area that could be expanded on is the edge detection. As explained in the background, today computer vision is still far from perfect. However that will hopefully change in the near future and things like Mechanical Turk can be dropped for a more efficient system that produces little to no error and works autonomously.
